---
title    : "Data visualisation, interactive data analysis, statistical programming"
author   : Garth Tarr
framework   : minimal
url      : {lib: "../../libraries", assets: "../../assets"}
mode     : selfcontained # {standalone, draft}
hitheme     : solarized_light
lecnum   : "01"
widgets  : [mathjax]
github      : {user: garthtarr, repo: visR, branch: gh-pages, name: Garth Tarr}
---

```{r Lab1, echo=FALSE, cache=FALSE}
#options(width = 65)
#library(knitr)
#library(knitcitations)
#bib <- read.bibtex('../../assets/msR.bib')
opts_chunk$set(tidy = TRUE, cache=FALSE) 
# tidy = TRUE makes the R code more readable (proper spacing)
# results = 'hide' is used for the actual version
# results = 'hold' can be used to ensure everything is working as expected
# and can also be used to provide "solutions" to our demonstrators
```

## Characterizing mutation-expression networks in multiple cancers


- <a href="https://twitter.com/shazanfar">Shila Ghazanfar's</a>  [BIS2015 poster](https://github.com/garthtarr/visR/blob/gh-pages/labs/02/BIS2015FF_Ghazanfar.pdf) has an accompanying Shiny app.
- 4,443 tumours analysed over 19 cancers from TCGA
- R Shiny app to explore results among different parameters is available to download at [www.maths.usyd.edu.au/u/sheilag/pacmen.zip](http://www.maths.usyd.edu.au/u/sheilag/pacmen.zip)

### Instructions

- Check out Shila's fast forward poster session slide <a href="https://github.com/garthtarr/visR/blob/gh-pages/labs/02/BIS2015FF_Ghazanfar.pdf"><i class="fa fa-link"></i></a>
- Download and unzip from [www.maths.usyd.edu.au/u/sheilag/pacmen.zip](http://www.maths.usyd.edu.au/u/sheilag/pacmen.zip)
- Install a few R packages including customised version of D3network using:

```{r, eval=FALSE}
install.packages(c("shiny","igraph","WGCNA","devtools"))
source("http://bioconductor.org/biocLite.R")
biocLite("impute")
devtools::install_github("garthtarr/networkD3")
```

- Run the app using a command like (replace `~/Downloads/` with the path to the `pacmen` unzipped folder):

```{r, eval=FALSE}
shiny::runApp('~/Downloads/pacmen')
```

- Alternatively you can run the app by opening either the `server.R` or the `ui.R` files in RStudio and then clicking the `Run App` button in the top right of the code editor window.

- Explore the app to get a feel for the sorts of things you can do with shiny.
- Check out Shila's twitter account <a href="https://twitter.com/shazanfar"><i class="fa fa-twitter"></i></a>

## Common issues

- It may take a little while for the app to load, be patient!  Serious computations are in progress.
- If the linkages are missing in the interactive graph, but you can see them in the static graph, you need to install a customised version of `networkD3`:

```{r, eval = FALSE}
devtools::install_github("garthtarr/networkD3")
```



## Reference

- 
